---
title: "PADDI enrichment analysis - Reactome"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Let's do some enrichment analysis.

```{r,libraries}

suppressPackageStartupMessages({
  library("gplots")
  library("reshape2")
  library("WGCNA")
  library("dplyr")
  library("DESeq2")
  library("mitch")
  library("MASS")
  library("kableExtra")
  library("ggplot2")
  library("eulerr")
  library("DT")
  library("xlsx")
})

load("qc.Rds")

```

## Load REACTOME gene sets

```{r,geneset}

reactome <- mitch::gmt_import("ReactomePathways_30oct24.gmt")

```

## Gene table

```{r,genetable}

gt <- read.table("../ref/gencode.v38.genetable.tsv")

rownames(gt) <- paste(gt[,1],gt[,2])

gt[,1] <- rownames(gt)

```

## Individual contrast analysis unstratified

Start with Reactome gene sets

Firstly we do the unstratified contrasts.

* crp_t0_adj
* crp_eos_adj
* crp_pod1_adj
* dex_t0_adj
* dex_eos_adj
* dex_pod1_adj

```{r,mitch1}

de <- crp_t0_adj
myname <- "crp_t0_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_crp_t0_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_eos_adj
myname <- "crp_eos_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_crp_eos_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_pod1_adj
myname <- "crp_pod1_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_crp_pod1_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

## dex
de <- dex_t0_adj
myname <- "dex_t0_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_dex_t0_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_eos_adj
myname <- "dex_eos_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_dex_eos_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_pod1_adj
myname <- "dex_pod1_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
m_dex_pod1_adj <- mres
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)




```

Now take a look at the interactions.

* m_crp_t0_adj
* m_crp_eos_adj
* m_crp_pod1_adj
* m_dex_t0_adj
* m_dex_eos_adj
* m_dex_pod1_adj


```{r,interactions}

m_crp_t0_adj_up <- subset(m_crp_t0_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_crp_t0_adj_dn <- subset(m_crp_t0_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

m_crp_eos_adj_up <- subset(m_crp_eos_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_crp_eos_adj_dn <- subset(m_crp_eos_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

m_crp_pod1_adj_up <- subset(m_crp_pod1_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_crp_pod1_adj_dn <- subset(m_crp_pod1_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

m_dex_t0_adj_up <- subset(m_dex_t0_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_dex_t0_adj_dn <- subset(m_dex_t0_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

m_dex_eos_adj_up <- subset(m_dex_eos_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_dex_eos_adj_dn <- subset(m_dex_eos_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

m_dex_pod1_adj_up <- subset(m_dex_pod1_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist>0)$set
m_dex_pod1_adj_dn <- subset(m_dex_pod1_adj$enrichment_result,p.adjustANOVA<0.01 & s.dist<0)$set

crp_l <- list("t0 CRP up"=m_crp_t0_adj_up,"t0 CRP dn"=m_crp_t0_adj_dn,
  "EOS CRP up"=m_crp_eos_adj_up, "EOS CRP dn"=m_crp_eos_adj_dn,
  "POD1 CRP up"=m_crp_pod1_adj_up, "POD1 CRP dn"=m_crp_pod1_adj_dn)

plot(euler(crp_l),quantities = TRUE)

crp_l <- list("EOS CRP up"=m_crp_eos_adj_up, "EOS CRP dn"=m_crp_eos_adj_dn,
  "POD1 CRP up"=m_crp_pod1_adj_up, "POD1 CRP dn"=m_crp_pod1_adj_dn)

plot(euler(crp_l),quantities = TRUE)

dex_l <- list("t0 dex up"=m_dex_t0_adj_up,"t0 dex dn"=m_dex_t0_adj_dn,
  "EOS dex up"=m_dex_eos_adj_up, "EOS dex dn"=m_dex_eos_adj_dn,
  "POD1 dex up"=m_dex_pod1_adj_up, "POD1 dex dn"=m_dex_pod1_adj_dn)

plot(euler(dex_l),quantities = TRUE)

dex_l <- list("EOS dex up"=m_dex_eos_adj_up, "EOS dex dn"=m_dex_eos_adj_dn,
  "POD1 dex up"=m_dex_pod1_adj_up, "POD1 dex dn"=m_dex_pod1_adj_dn)

plot(euler(dex_l),quantities = TRUE)


cross_eos <- list("EOS dex up"=m_dex_eos_adj_up, "EOS dex dn"=m_dex_eos_adj_dn,
  "EOS CRP up"=m_crp_eos_adj_up, "EOS CRP dn"=m_crp_eos_adj_dn)

plot(euler(cross_eos),quantities = TRUE)


cross_pod1 <- list("POD1 dex up"=m_dex_pod1_adj_up, "POD1 dex dn"=m_dex_pod1_adj_dn,
  "POD1 CRP up"=m_crp_pod1_adj_up, "POD1 CRP dn"=m_crp_pod1_adj_dn)

plot(euler(cross_pod1),quantities = TRUE)

intersect(m_crp_pod1_adj_up, m_dex_pod1_adj_dn)

intersect(m_crp_pod1_adj_dn, m_dex_pod1_adj_up)


```

## Multi-contrast enrichment analysis.

```{r,multi1}

l1 <- list("crp_t0_adj"=crp_t0_adj,"crp_eos_adj"=crp_eos_adj,
  "crp_pod1_adj"=crp_pod1_adj,"dex_t0_adj"=dex_t0_adj,
  "dex_eos_adj"=dex_eos_adj,"dex_pod1_adj"=dex_pod1_adj)

m1 <- mitch_import(x=l1, DEtype="deseq2", geneTable=gt )

mm1 <- mitch_calc(x=m1,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm1$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:9)]
rownames(top) <- top[,1]
top[,1] = NULL

colfunc <- colorRampPalette(c("blue", "white", "red"))

heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none",
    margins = c(6,25), cexRow=0.6, cexCol=0.8 )

as.matrix(top) |> kbl(caption="Top REACTOMEs in multi enrichment analysis") |> kable_paper("hover", full_width = F)

mitch_report(res=mm1,outfile="multireactome_all_mitchreport.html",overwrite=TRUE)

```

This might work better if we work on each timepoint separately.

```{r,multi1_t0}

l1 <- list("crp_t0_adj"=crp_t0_adj, "dex_t0_adj"=dex_t0_adj)
m1 <- mitch_import(x=l1, DEtype="deseq2", geneTable=gt )
mm1 <- mitch_calc(x=m1,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm1$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:5)]
rownames(top) <- top[,1]
top[,1] = NULL
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none", margins = c(6,25), cexRow=0.6, cexCol=0.8 )
as.matrix(top) |> kbl(caption="Top REACTOMEs in multi enrichment analysis at t0") |> kable_paper("hover", full_width = F)
mitch_report(res=mm1,outfile="multireactome_t0_mitchreport.html",overwrite=TRUE)

l1 <- list("crp_eos_adj"=crp_eos_adj, "dex_eos_adj"=dex_eos_adj)
m1 <- mitch_import(x=l1, DEtype="deseq2", geneTable=gt )
mm1 <- mitch_calc(x=m1,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm1$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:5)]
rownames(top) <- top[,1]
top[,1] = NULL
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none", margins = c(6,25), cexRow=0.6, cexCol=0.8 )
as.matrix(top) |> kbl(caption="Top REACTOMEs in multi enrichment analysis at EOS") |> kable_paper("hover", full_width = F)
mitch_report(res=mm1,outfile="multireactome_eos_mitchreport.html",overwrite=TRUE)

l1 <- list("crp_pod1_adj"=crp_pod1_adj, "dex_pod1_adj"=dex_pod1_adj)
m1 <- mitch_import(x=l1, DEtype="deseq2", geneTable=gt )
mm1 <- mitch_calc(x=m1,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm1$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:5)]
rownames(top) <- top[,1]
top[,1] = NULL
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none", margins = c(6,25), cexRow=0.6, cexCol=0.8 )
as.matrix(top) |> kbl(caption="Top REACTOMEs in multi enrichment analysis at POD1") |> kable_paper("hover", full_width = F)
mitch_report(res=mm1,outfile="multireactome_pod1_mitchreport.html",overwrite=TRUE)

```


## Individual contrast analysis stratified

* crp_t0_a_adj
* crp_t0_b_adj
* crp_eos_a_adj
* crp_eos_b_adj
* crp_pod1_a_adj
* crp_pod1_b_adj
* dex_crplo_t0_adj
* dex_crphi_t0_adj
* dex_crplo_eos_adj
* dex_crphi_eos_adj
* dex_crplo_pod1_adj
* dex_crphi_pod1_adj

```{r,mitch2}

de <- crp_t0_a_adj
myname <- "crp_t0_a_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_t0_b_adj
myname <- "crp_t0_b_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_eos_a_adj
myname <- "crp_eos_a_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_eos_b_adj
myname <- "crp_eos_b_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_pod1_a_adj
myname <- "crp_pod1_a_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- crp_pod1_b_adj
myname <- "crp_pod1_b_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crplo_t0_adj
myname <- "dex_crplo_t0_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crphi_t0_adj
myname <- "dex_crphi_t0_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crplo_eos_adj
myname <- "dex_crplo_eos_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crphi_eos_adj
myname <- "dex_crphi_eos_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crplo_pod1_adj
myname <- "dex_crplo_pod1_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

de <- dex_crphi_pod1_adj
myname <- "dex_crphi_pod1_adj"

m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
mres <- mitch_calc(x=m,genesets=reactome,minsetsize=5,cores=16,priority="effect")
mtop <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05),30)
mtop |> kbl(caption=paste(myname,"REACTOME")) |> kable_paper("hover", full_width = F)
vec <- mtop$s.dist
names(vec) <- mtop$set
vec <- sort(vec)
par( mar = c(5.1, 25.1, 4.1, 2.1) )
barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=myname)
par( mar = c(5.1, 4.1, 4.1, 2.1) )
mitch_report(res=mres,outfile=paste(myname,"_mitchreport.html",sep=""),overwrite=TRUE)

```


Multi-contrast enrichment anslysis.

```{r,multi2}

l2 <- list("crp_t0_a"=crp_t0_a_adj,
"crp t0 b"=crp_t0_b_adj,
"crp eos a"=crp_eos_a_adj,
"crp eos b"=crp_eos_b_adj,
"crp _pod1 a"=crp_pod1_a_adj,
"crp pod1 b"=crp_pod1_b_adj,
"dex crplo t0"=dex_crplo_t0_adj,
"dex crphi t0"=dex_crphi_t0_adj,
"dex crplo eos"=dex_crplo_eos_adj,
"dex crphi eos"=dex_crphi_eos_adj,
"dex crplo pod1"=dex_crplo_pod1_adj,
"dex crphi pod1"=dex_crphi_pod1_adj)

m2 <- mitch_import(x=l2, DEtype="deseq2", geneTable=gt )

mm2 <- mitch_calc(x=m2,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm2$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:15)]
rownames(top) <- top[,1]
top[,1] = NULL
colnames(top) <- gsub("^s\\.","",colnames(top))

colfunc <- colorRampPalette(c("blue", "white", "red"))

heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none",
    margins = c(7,20), cexRow=0.6, cexCol=0.7 )

as.matrix(top) |>
  kbl(caption="Top REACTOMEs in multi enrichment analysis") |>
  kable_paper("hover", full_width = F)

mitch_report(res=mm2,outfile="multireactomestratified_all_mitchreport.html",overwrite=TRUE)

```

This might work better if we work on each timepoint separately.

```{r,multi2_t0}

l2 <- list("crp_t0_a"=crp_t0_a_adj, "crp t0 b"=crp_t0_b_adj,
  "dex crplo t0"=dex_crplo_t0_adj, "dex crphi t0"=dex_crphi_t0_adj)
m2 <- mitch_import(x=l2, DEtype="deseq2", geneTable=gt )
mm2 <- mitch_calc(x=m2,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm2$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:7)]
rownames(top) <- top[,1]
top[,1] = NULL
colnames(top) <- gsub("^s\\.","",colnames(top))
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none",
    margins = c(7,20), cexRow=0.6, cexCol=0.7 )
as.matrix(top) |>
  kbl(caption="Top REACTOMEs in multi enrichment analysis stratified t0") |>
  kable_paper("hover", full_width = F)
#mitch_report(res=mm2,outfile="multireactomestratified_t0_mitchreport.html",overwrite=TRUE)

l2 <- list("crp_eos_a"=crp_eos_a_adj, "crp eos b"=crp_eos_b_adj,
  "dex crplo eos"=dex_crplo_eos_adj, "dex crphi eos"=dex_crphi_eos_adj)
m2 <- mitch_import(x=l2, DEtype="deseq2", geneTable=gt )
mm2 <- mitch_calc(x=m2,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm2$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:7)]
rownames(top) <- top[,1]
top[,1] = NULL
colnames(top) <- gsub("^s\\.","",colnames(top))
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none",
    margins = c(7,20), cexRow=0.6, cexCol=0.7 )
as.matrix(top) |>
  kbl(caption="Top REACTOMEs in multi enrichment analysis stratified EOS") |>
  kable_paper("hover", full_width = F)
#mitch_report(res=mm2,outfile="multireactomestratified_eos_mitchreport.html",overwrite=TRUE)

l2 <- list("crp_pod1_a"=crp_pod1_a_adj, "crp pod1 b"=crp_pod1_b_adj,
  "dex crplo pod1"=dex_crplo_pod1_adj, "dex crphi pod1"=dex_crphi_pod1_adj)
m2 <- mitch_import(x=l2, DEtype="deseq2", geneTable=gt )
mm2 <- mitch_calc(x=m2,genesets=reactome,minsetsize=5,cores=16,priority="effect")
top <- head(subset (mm2$enrichment_result,p.adjustMANOVA<0.05),50)
top <- top[,c(1,4:7)]
rownames(top) <- top[,1]
top[,1] = NULL
colnames(top) <- gsub("^s\\.","",colnames(top))
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(as.matrix(top),trace="none",col=colfunc(25),scale="none",
    margins = c(7,20), cexRow=0.6, cexCol=0.7 )
as.matrix(top) |>
  kbl(caption="Top REACTOMEs in multi enrichment analysis stratified POD1") |>
  kable_paper("hover", full_width = F)
#mitch_report(res=mm2,outfile="multireactomestratified_pod1_mitchreport.html",overwrite=TRUE)

```

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
