---
title: "PADDI RNA expression downstream analysis"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

The differential analysis was already conducted using the qc.Rmd script.
Here we are doing downstream analysis including pathway analysis and charts.

```{r,libraries}

suppressPackageStartupMessages({
  library("gplots")
  library("reshape2")
  library("WGCNA")
  library("dplyr")
  library("DESeq2")
  library("mitch")
  library("MASS")
  library("kableExtra")
  library("ggplot2")
  library("eulerr")
})

load("qc.Rds")

```

## Functions

```{r,myfunctions}

volcanoplot <- function(x,title) {
  tot <- nrow(x)
  sig <- subset(x,padj<0.05)
  nsig <- nrow(subset(x,padj<0.05))
  dn <- nrow(subset(x,padj<0.05 & log2FoldChange<0))
  up <- nrow(subset(x,padj<0.05 & log2FoldChange>0))
  header <- paste(title,":", tot, "genes,", nsig ,"@5% FDR,", up, "up,", dn, "dn")

  plot(x$log2FoldChange,-log10(x$pvalue),bty="n",
    cex=0.5, pch=19, xlab="log2 fold change",
    ylab="-log10 p-value", main=header)

  points(sig$log2FoldChange,-log10(sig$pvalue),
    cex=0.5, pch=19, col="red")
}

smearplot <- function(x,title) {
  tot <- nrow(x)
  sig <- subset(x,padj<0.05)
  nsig <- nrow(subset(x,padj<0.05))
  dn <- nrow(subset(x,padj<0.05 & log2FoldChange<0))
  up <- nrow(subset(x,padj<0.05 & log2FoldChange>0))
  header <- paste(title,":", tot, "genes,", nsig ,"@5% FDR,", up, "up,", dn, "dn")

  plot(log10(x$baseMean),x$log2FoldChange,bty="n",
    cex=0.5, pch=19, xlab="log10 base mean",
    ylab="log2 fold change", main=header)

  points(log10(sig$baseMean),sig$log2FoldChange,
    cex=0.5, pch=19, col="red")
}

debar <- function(l) {
  nn <- sapply(l,function(x) {
    nup <- nrow(subset(x,padj<0.05 & log2FoldChange>0 ))
    ndn <- nrow(subset(x,padj<0.05 & log2FoldChange<0 )) *-1
    return(c(nup,ndn))
  })
  rownames(nn) <- c("up","dn")
  par(mar = c(8.1, 4.1, 4.1, 2.1))
  barplot(nn[1,],col="blue",beside = TRUE, border=TRUE,ylim=c(min(nn),max(nn)) , las=2)
  barplot(nn[2,],add=TRUE,col="red",names.arg = "", ylab="", yaxt="n")
  mtext("blue=up,red=down")
  par(mar = c(5.1, 4.1, 4.1, 2.1))
}

myeuler <- function(l) {
  ups <- lapply(l,function(x) { rownames(subset(x,padj<0.05 & log2FoldChange>0) )})
  dns <- lapply(l,function(x) { rownames(subset(x,padj<0.05 & log2FoldChange<0) )})
  names(ups) <- paste(names(ups),"up")
  names(dns) <- paste(names(dns),"dn")
  ll <- c(ups,dns)
  plot(euler(ll),quantities = TRUE)
}

```

## DE results

### Unadjusted for clinical covariates

#### Unstratified, unadjusted

* crp_t0
* crp_eos
* crp_pod1
* avb_t0
* avb_eos
* avb_pod1

```{r,l1_basic_charts}

l1 <- list("crp_t0"=crp_t0, "crp_eos"=crp_eos, "crp_pod1"=crp_pod1,
  "avb_t0"=avb_t0,"avb_eos"=avb_eos,"avb_pod1"=avb_pod1 )

lapply(1:length(l1),function(i) {
  volcanoplot(x=l1[[i]], title=names(l1)[[i]])
  smearplot(x=l1[[i]], title=names(l1)[[i]])
  head(x,50) |>
    kbl(caption = paste(names(l1)[[i]],"top 50 genes")) |>
    kable_paper("hover", full_width = F)
} )

debar(l1) ; par(mar = c(5.1, 4.1, 4.1, 2.1))

myeuler(list("crp_t0"=crp_t0,"avb_t0"=avb_t0))
myeuler(list("crp_eos"=crp_eos,"avb_eos"=avb_eos))
myeuler(list("crp_pod1"=crp_pod1,"avb_pod1"=avb_pod1))

```

#### Stratified, unadjusted

* crp_t0_a
* crp_t0_b
* crp_eos_a
* crp_eos_b
* crp_pod1_a
* crp_pod1_b
* avb_crplo_t0
* avb_crphi_t0
* avb_crplo_eos
* avb_crphi_eos
* avb_crplo_pod1
* avb_crphi_pod1

```{r,l2_basic_charts}

l2 <- list("crp_t0_a"=crp_t0_a, "crp_t0_b"=crp_t0_b,
  "crp_eos_a"=crp_eos_a, "crp_eos_b"=crp_eos_b,
  "crp_pod1_a"= crp_pod1_a, "crp_pod1_b"=crp_pod1_b,
  "avb_crplo_t0"=avb_crplo_t0, "avb_crphi_t0"= avb_crphi_t0,
  "avb_crplo_eos"=avb_crplo_eos, "avb_crphi_eos"=avb_crphi_eos,
  "avb_crplo_pod1"=avb_crplo_pod1, "avb_crphi_pod1"=avb_crphi_pod1)

lapply(1:length(l2),function(i) {
  volcanoplot(x=l2[[i]], title=names(l2)[[i]])
  smearplot(x=l2[[i]], title=names(l2)[[i]])
  head(x,50) |>
    kbl(caption = paste(names(l2)[[i]],"top 50 genes")) |>
    kable_paper("hover", full_width = F)
} )

debar(l2) ; par(mar = c(5.1, 4.1, 4.1, 2.1))

myeuler(list("crp_t0_a"=crp_t0_a, "crp_t0_b"=crp_t0_b,"avb_crplo_t0"=avb_crplo_t0, "avb_crphi_t0"= avb_crphi_t0))
myeuler(list("crp_eos_a"=crp_eos_a, "crp_eos_b"=crp_eos_b,"avb_crplo_eos"=avb_crplo_eos, "avb_crphi_eos"=avb_crphi_eos))
myeuler(list("crp_pod1_a"= crp_pod1_a, "crp_pod1_b"=crp_pod1_b,"avb_crplo_pod1"=avb_crplo_pod1, "avb_crphi_pod1"=avb_crphi_pod1))

```

### Adjusted for clinical covariates

#### Unstratified, adjusted

* crp_t0_adj
* crp_eos_adj
* crp_pod1_adj
* avb_t0_adj
* avb_eos_adj
* avb_pod1_adj

```{r,l3_basic_charts}

l3 <- list("crp_t0_adj"=crp_t0_adj, "crp_eos_adj"=crp_eos_adj, "crp_pod1_adj"=crp_pod1_adj,
  "avb_t0_adj"=avb_t0_adj,"avb_eos_adj"=avb_eos_adj,"avb_pod1_adj"=avb_pod1_adj )

lapply(1:length(l3),function(i) {
  volcanoplot(x=l3[[i]], title=names(l3)[[i]])
  smearplot(x=l3[[i]], title=names(l3)[[i]])
  head(x,50) |>
    kbl(caption = paste(names(l3)[[i]],"top 50 genes")) |>
    kable_paper("hover", full_width = F)
} )

debar(l3) ; par(mar = c(5.1, 4.1, 4.1, 2.1))

myeuler(list("crp_t0_adj"=crp_t0_adj,"avb_t0_adj"=avb_t0_adj))
myeuler(list("crp_eos_adj"=crp_eos_adj,"avb_eos_adj"=avb_eos_adj))
myeuler(list("crp_pod1_adj"=crp_pod1_adj,"avb_pod1_adj"=avb_pod1_adj))

```

#### Stratified, adjusted

* crp_t0_a_adj
* crp_t0_b_adj
* crp_eos_a_adj
* crp_eos_b_adj
* crp_pod1_a_adj
* crp_pod1_b_adj
* avb_crplo_t0_adj
* avb_crphi_t0_adj
* avb_crplo_eos_adj
* avb_crphi_eos_adj
* avb_crplo_pod1_adj
* avb_crphi_pod1_adj

```{r,l4_basic_charts}

l4 <- list("crp_t0_a_adj"=crp_t0_a_adj, "crp_t0_b_adj"=crp_t0_b_adj,
  "crp_eos_a_adj"=crp_eos_a_adj, "crp_eos_b_adj"=crp_eos_b_adj,
  "crp_pod1_a_adj"= crp_pod1_a_adj, "crp_pod1_b_adj"=crp_pod1_b_adj,
  "avb_crplo_t0_adj"=avb_crplo_t0_adj, "avb_crphi_t0_adj"= avb_crphi_t0_adj,
  "avb_crplo_eos_adj"=avb_crplo_eos_adj, "avb_crphi_eos_adj"=avb_crphi_eos_adj,
  "avb_crplo_pod1_adj"=avb_crplo_pod1_adj, "avb_crphi_pod1_adj"=avb_crphi_pod1_adj)

lapply(1:length(l4),function(i) {
  volcanoplot(x=l2[[i]], title=names(l2)[[i]])
  smearplot(x=l2[[i]], title=names(l2)[[i]])
  head(x,50) |>
    kbl(caption = paste(names(l2)[[i]],"top 50 genes")) |>
    kable_paper("hover", full_width = F)
} )

debar(l4) ; par(mar = c(5.1, 4.1, 4.1, 2.1))

myeuler(list("crp_t0_a_adj"=crp_t0_a_adj, "crp_t0_b_adj"=crp_t0_b_adj,
  "avb_crplo_t0_adj"=avb_crplo_t0_adj, "avb_crphi_t0_adj"= avb_crphi_t0_adj))

myeuler(list("crp_eos_a_adj"=crp_eos_a_adj, "crp_eos_b_adj"=crp_eos_b_adj,
  "avb_crplo_eos_adj"=avb_crplo_eos_adj, "avb_crphi_eos_adj"=avb_crphi_eos_adj))

myeuler(list("crp_pod1_a_adj"= crp_pod1_a_adj, "crp_pod1_b_adj"=crp_pod1_b_adj,
  "avb_crplo_pod1_adj"=avb_crplo_pod1_adj, "avb_crphi_pod1_adj"=avb_crphi_pod1_adj))

```
## Session information

For reproducibility

```{r,session}

sessionInfo()

```

