---
title: "PADDI RNA expression analysis"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Let's QC this data.

```{r,libraries}

library("gplots")
library("reshape2")

```

## Multi-qc results

Please have a look at the multiQC report. 
Here are a few key points:

* Skewer trimming resulted in loss of only a tiny number of bases.
This indicates the sequence quality is very high.

* Fastqc results showing the number of unique and duplicate reads indicates a few samples with <10M unique reads.

* Per seqence GC content showed an unusual profile for two samples.
PG1423-EOS R1 and R2 had GC profile max at 40% compared to the mean.
PG2090-EOS also showed an unusual pattern with underrepresented low GC%.

* Sequence duplication levels were elevated for some fastq files.
Here are the files of concern, with <20% unique reads:
PG3627-POD1_S86_R1_001
PG3627-POD1_S86_R2_001
PG3609-T0_S317_R1_001
PG2090-EOS_S134_R1_001
PG2090-EOS_S134_R2_001

* There were two files with overrepresented sequences: PG2090-EOS R1 and R2. Others are okay.

* Adapter content was very low which is good.

The fastq files were also checked with validatefastq-assembly which looks for signs of file corruption which can occur in large data transfers.
No problematic files were detected.

## rRNA amount

Ribosomal RNA carryover can be a source of noise.
The proportion should be <10% and there were a few samples in excess of this including PG2020-EOS, PG815-EOS, PG1452-EOS and PG702-POD1.
 
```{r,rrna}

rrna <- read.table("rrna_stats.txt")
rrna <- rrna[,c(1,5)]
rrna$V1 <- sapply(strsplit(rrna$V1,"\\."),"[[",1)
rrna$V5 <- gsub("\\(","",rrna$V5)
rrna$V5 <- gsub("%","",rrna$V5)
rrna$V5 <- as.numeric(rrna$V5)
str(rrna)

rrna2 <- rrna[,2]
names(rrna2) <- rrna[,1]

par(mar=c(5,8,3,1))
barplot(rrna2,horiz=TRUE,las=1,cex.names=0.5,main="rRNA carryover")
rrna2 <- rrna2[order(-rrna2)]
barplot(head(rrna2,20),horiz=TRUE,las=1,cex.names=0.6,main="rRNA carryover")

```

## Load the data

```{r,load}

tmp <- read.table("3col.tsv.gz",header=FALSE)
x <- as.matrix(acast(tmp, V2~V1, value.var="V3", fun.aggregate = sum))
x <- as.data.frame(x)
accession <- sapply((strsplit(rownames(x),"\\|")),"[[",2)
symbol<-sapply((strsplit(rownames(x),"\\|")),"[[",6)
x$geneid <- paste(accession,symbol)
xx <- aggregate(. ~ geneid,x,sum)
rownames(xx) <- xx$geneid
colnames <- gsub("T0R","T0",colnames(xx))
xx$geneid = NULL
xx <- round(xx)
xx[1:10,1:6]

```

## Number of reads per sample

Let's look at the number of reads per sample

Most samples were in the range of 25-30 million assigned reads.
Just 2 samples had less than 20 million reads: PG1452-EOS and PG1423-EOS.
The maximum read count was about 40 million for PG7072-EOS.

```{r,numreads}

xxcs <- colSums(xx)
par(mar=c(5,8,3,1))
barplot(xxcs,horiz=TRUE,las=1,main="no. reads per sample")
barplot(head(xxcs[order(xxcs)],20),horiz=TRUE,las=1,main="lowest no. reads per sample")
barplot(head(xxcs[order(-xxcs)],20),horiz=TRUE,las=1,main="highest no. reads per sample")

```


## MDS

Some outliers are apparent.

PG2090-EOS to the left of the chart - this is clearly the effect of rRNA carryover.
Other samples over to the left of the chart include PG815-EOS, PG145-EOS and PG702-POD1 which all have elevated rRNA.

```{r,mds1}

mds <- cmdscale(dist(t(xx)))

par(mar=c(5,5,3,1))
minx <- min(mds[,1])
maxx <- max(mds[,1])
miny <- min(mds[,2])
maxy <- max(mds[,2])

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2", 
  xlim=c(minx*1.1,maxx*1.1), ylim = c(miny*1.1,maxy*1.1) ,
  type = "p", col="gray", pch=19, cex.axis=1.3,cex.lab=1.3, bty='n')
text(mds, labels=rownames(mds), cex=0.8) 

col <- rownames(mds)
col <- sapply(strsplit(col,"-"),"[[",2)
col <- gsub("T0","lightblue",col)
col <- gsub("POD1","orange",col)
col <- gsub("EOS","pink",col)

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2", 
  xlim=c(minx*1.1,maxx*1.1), ylim = c(miny*1.1,maxy*1.1) , cex=1.5 , 
  type = "p", col=col, pch=19, cex.axis=1.3,cex.lab=1.3, bty='n')
#text(mds, labels=rownames(mds), cex=0.8) 
mtext("blue=T0, orange=POD1, pink=EOS")

```

Exclude PG2090-EOS and repeat the analysis.

```{r,mds2}

xx <- xx[,grep("PG2090-EOS",colnames(xx),invert=TRUE)]

mds <- cmdscale(dist(t(xx)))

par(mar=c(5,5,3,1))
minx <- min(mds[,1])
maxx <- max(mds[,1])
miny <- min(mds[,2])
maxy <- max(mds[,2])

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2",
  xlim=c(minx*1.1,maxx*1.1), ylim = c(miny*1.1,maxy*1.1) ,
  type = "p", col="gray", pch=19, cex.axis=1.3,cex.lab=1.3, bty='n')
text(mds, labels=rownames(mds), cex=0.8)

col <- rownames(mds)
col <- sapply(strsplit(col,"-"),"[[",2)
col <- gsub("T0","lightblue",col)
col <- gsub("POD1","orange",col)
col <- gsub("EOS","pink",col)

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2",
  xlim=c(minx*1.1,maxx*1.1), ylim = c(miny*1.1,maxy*1.1) , cex=1.5 ,
  type = "p", col=col, pch=19, cex.axis=1.3,cex.lab=1.3, bty='n')
#text(mds, labels=rownames(mds), cex=0.8) 
mtext("blue=T0, orange=POD1, pink=EOS")

```

In the MDS plot with PG2090-EOS removed, there appears to be some separation of T0, POD1 and EOS samples.
POD1 (orange) are more towards the upper side of the chart and T0 (blue) are toward the bottom right.
EOS (pink) are quite spread out.

## Conclusion

PG2090-EOS suffered rRNA carryover and needs to be re-prepared. 
The other samples with slightly higher rRNA are not a problem as the rRNA can be corrected for statistically.
not sure what to do about samples with low numbers of unique reads.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
