---
title: "PADDI RNA expression analysis - Timecourse Analysis - Downstream"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Time course analysis of PADDI genomics data.

```{r,libraries}

suppressPackageStartupMessages({
  library("gplots")
  library("reshape2")
  library("WGCNA")
  library("dplyr")
  library("DESeq2")
  library("mitch")
  library("MASS")
  library("eulerr")
})

```

## Load the data

```{r,load}

load("tca_pairwise.Rdata")

```

## Overview

We will run timecourse analysis with a simple pairwise approach.
For each of the groups below, this will involve three comparisons:

* T0 vs EOS

* EOS vs POD1

* T0 vs POD1

The groups/subgroups we will look at are:

1. Timecourse in low CRP group

2. Timecourse in high CRP group

3. Timecourse in low CRP group and treatment group A

4. Timecourse in low CRP group and treatment group B

5. Timecourse in high CRP group and treatment group A

6. Timecourse in high CRP group and treatment group B

In this report I will be loading in the DESeq2 objects and conducting downstream analysis.

## Functions

```{r,decompfunc}

dequant1 <- function(x) {
  all <- nrow(x)
  nde <- length(rownames(subset(x,padj<0.05)))
  nup <- length(rownames(subset(x,padj<0.05 & log2FoldChange>0)))
  ndn <- length(rownames(subset(x,padj<0.05 & log2FoldChange<0)))
  return(c("all"=all,"nde"=nde,
    "up"=nup,"dn"=ndn))
}

dequant2 <- function(x,y) {
  xup <- rownames(subset(x,padj<0.05 & log2FoldChange>0))
  xdn <- rownames(subset(x,padj<0.05 & log2FoldChange<0))
  yup <- rownames(subset(y,padj<0.05 & log2FoldChange>0))
  ydn <- rownames(subset(y,padj<0.05 & log2FoldChange<0))
  nxup <- length(xup)
  nxdn <- length(xdn)
  nyup <- length(yup)
  nydn <- length(ydn)
  iup <- intersect(xup,yup)
  oxup <- setdiff(xup,iup)
  oyup <- setdiff(yup,iup)
  unup <- union(xup,yup)
  idn <- intersect(xdn,ydn)
  oxdn <- setdiff(xdn,idn)
  oydn <- setdiff(ydn,idn)
  undn <- union(xdn,ydn)
  myintersect <- length(c(iup,idn))
  myunion <- length(c(unup,undn))
  myjac <- myintersect / myunion

  res <- c("xup"=nxup,"xdn"=nxdn,"yup"=nyup,"ydn"=nydn,
    "intersect"=myintersect,"union"=myunion,"jac"=myjac)
  return(res)
}

de_euler <- function(x,y,xname,yname) {

  xup <- rownames(subset(x,padj<0.05 & log2FoldChange>0))
  xdn <- rownames(subset(x,padj<0.05 & log2FoldChange<0))
  yup <- rownames(subset(y,padj<0.05 & log2FoldChange>0))
  ydn <- rownames(subset(y,padj<0.05 & log2FoldChange<0))

  xname_up <- paste(xname,"up")
  xname_dn <- paste(xname,"dn")
  yname_up <- paste(yname,"up")
  yname_dn <- paste(yname,"dn")

  v1 <- list(xup, xdn, yup, ydn)
  names(v1) <- c(xname_up, xname_dn, yname_up, yname_dn)
  plot(euler(v1),quantities = TRUE)

}

```

### Object names

#### Timecourse in low CRP group

tc_lo_t0veos
tc_lo_eosvpod1
tc_lo_t0vpod1

tc_lo_t0veos_adj
tc_lo_eosvpod1_adj
tc_lo_t0vpod1_adj

```{r,quant1}

lapply(list("tc_lo_t0veos"=tc_lo_t0veos,
  "tc_lo_eosvpod1"=tc_lo_eosvpod1,
  "tc_lo_t0vpod1"=tc_lo_t0vpod1,
  "tc_lo_t0veos_adj"=tc_lo_t0veos_adj,
  "tc_lo_eosvpod1_adj"=tc_lo_eosvpod1_adj,
  "tc_lo_t0vpod1_adj"=tc_lo_t0vpod1_adj),dequant1)

```

#### Timecourse in high CRP group

tc_hi_t0veos
tc_hi_eosvpod1
tc_hi_t0vpod1

tc_hi_t0veos_adj
tc_hi_eosvpod1_adj
tc_hi_t0vpod1_adj

```{r,quant2}

lapply(list("tc_hi_t0veos"=tc_hi_t0veos,
  "tc_hi_eosvpod1"=tc_hi_eosvpod1,
  "tc_hi_t0vpod1"=tc_hi_t0vpod1,
  "tc_hi_t0veos_adj"=tc_hi_t0veos_adj,
  "tc_hi_eosvpod1_adj"=tc_hi_eosvpod1_adj,
  "tc_hi_t0vpod1_adj"=tc_hi_t0vpod1_adj),dequant1)

```

#### Timecourse in low CRP group and treatment group A

tc_lo_a_t0veos
tc_lo_a_eosvpod1
tc_lo_a_t0vpod1

tc_lo_a_t0veos_adj
tc_lo_a_eosvpod1_adj
tc_lo_a_t0vpod1_adj

```{r,quant3}

lapply(list("tc_lo_a_t0veos"=tc_lo_a_t0veos,
  "tc_lo_a_eosvpod1"=tc_lo_a_eosvpod1,
  "tc_lo_a_t0vpod1"=tc_lo_a_t0vpod1,
  "tc_lo_a_t0veos_adj"=tc_lo_a_t0veos_adj,
  "tc_lo_a_eosvpod1_adj"=tc_lo_a_eosvpod1_adj,
  "tc_lo_a_t0vpod1_adj"=tc_lo_a_t0vpod1_adj),dequant1)

```

#### Timecourse in low CRP group and treatment group B

tc_lo_b_t0veos
tc_lo_b_eosvpod1
tc_lo_b_t0vpod1

tc_lo_b_t0veos_adj
tc_lo_b_eosvpod1_adj
tc_lo_b_t0vpod1_adj

```{r,quant4}

lapply(list("tc_lo_b_t0veos"=tc_lo_b_t0veos,
  "tc_lo_b_eosvpod1"=tc_lo_b_eosvpod1,
  "tc_lo_b_t0vpod1"=tc_lo_b_t0vpod1,
  "tc_lo_b_t0veos_adj"=tc_lo_b_t0veos_adj,
  "tc_lo_b_eosvpod1_adj"=tc_lo_b_eosvpod1_adj,
  "tc_lo_b_t0vpod1_adj"=tc_lo_b_t0vpod1_adj),dequant1)

```


#### Timecourse in high CRP group and treatment group A

tc_hi_a_t0veos
tc_hi_a_eosvpod1
tc_hi_a_t0vpod1

tc_hi_a_t0veos_adj
tc_hi_a_eosvpod1_adj
tc_hi_a_t0vpod1_adj

```{r,quant5}

lapply(list("tc_hi_a_t0veos"=tc_hi_a_t0veos,
  "tc_hi_a_eosvpod1"=tc_hi_a_eosvpod1,
  "tc_hi_a_t0vpod1"=tc_hi_a_t0vpod1,
  "tc_hi_a_t0veos_adj"=tc_hi_a_t0veos_adj,
  "tc_hi_a_eosvpod1_adj"=tc_hi_a_eosvpod1_adj,
  "tc_hi_a_t0vpod1_adj"=tc_hi_a_t0vpod1_adj),dequant1)

```

#### Timecourse in high CRP group and treatment group B

tc_hi_b_t0veos
tc_hi_b_eosvpod1
tc_hi_b_t0vpod1

tc_hi_b_t0veos_adj
tc_hi_b_eosvpod1_adj
tc_hi_b_t0vpod1_adj

```{r,quant6}

lapply(list("tc_hi_b_t0veos"=tc_hi_b_t0veos,
  "tc_hi_b_eosvpod1"=tc_hi_b_eosvpod1,
  "tc_hi_b_t0vpod1"=tc_hi_b_t0vpod1,
  "tc_hi_b_t0veos_adj"=tc_hi_b_t0veos_adj,
  "tc_hi_b_eosvpod1_adj"=tc_hi_b_eosvpod1_adj,
  "tc_hi_b_t0vpod1_adj"=tc_hi_b_t0vpod1_adj),dequant1)

```
### Effect of correction for cell types

```{r,cellcorrection}

# low CRP group
dequant2(tc_lo_t0veos, tc_lo_t0veos_adj)
de_euler(x=tc_lo_t0veos, y=tc_lo_t0veos_adj, xname="tc_lo_t0veos",yname="tc_lo_t0veos_adj")

dequant2(tc_lo_eosvpod1, tc_lo_eosvpod1_adj)
de_euler(x=tc_lo_eosvpod1, y=tc_lo_eosvpod1_adj, xname="tc_lo_eosvpod1",yname="tc_lo_eosvpod1_adj")

dequant2(tc_lo_t0vpod1, tc_lo_t0vpod1_adj)
de_euler(x=tc_lo_t0vpod1, y=tc_lo_t0vpod1_adj, xname="tc_lo_t0vpod1", yname="tc_lo_t0vpod1_adj")

# high CRP group
dequant2(tc_hi_t0veos, tc_hi_t0veos_adj)
de_euler(x=tc_hi_t0veos, y=tc_hi_t0veos_adj, xname="tc_hi_t0veos", yname="tc_hi_t0veos_adj")

dequant2(tc_hi_eosvpod1, tc_hi_eosvpod1_adj)
de_euler(x=tc_hi_eosvpod1,y=tc_hi_eosvpod1_adj, xname="tc_hi_eosvpod1", yname="tc_hi_eosvpod1_adj")

dequant2(tc_hi_t0vpod1, tc_hi_t0vpod1_adj)
de_euler(x=tc_hi_t0vpod1, y=tc_hi_t0vpod1_adj, xname="tc_hi_t0vpod1", yname="tc_hi_t0vpod1_adj")

# Timecourse in low CRP group and treatment group A
dequant2(tc_lo_a_t0veos,tc_lo_a_t0veos_adj)
de_euler(x=tc_lo_a_t0veos, y=tc_lo_a_t0veos_adj, xname="tc_lo_a_t0veos", yname="tc_lo_a_t0veos_adj")

dequant2(tc_lo_a_eosvpod1, tc_lo_a_eosvpod1_adj)
de_euler(x=tc_lo_a_eosvpod1, y=tc_lo_a_eosvpod1_adj, xname="tc_lo_a_eosvpod1", yname="tc_lo_a_eosvpod1_adj")

dequant2(tc_lo_a_t0vpod1,tc_lo_a_t0vpod1_adj)
de_euler(x=tc_lo_a_t0vpod1, y=tc_lo_a_t0vpod1_adj, xname="tc_lo_a_t0vpod1", yname="tc_lo_a_t0vpod1_adj")

# Timecourse in low CRP group and treatment group B
dequant2(tc_lo_b_t0veos,tc_lo_b_t0veos_adj)
de_euler(x=tc_lo_b_t0veos, y=tc_lo_b_t0veos_adj, xname="tc_lo_b_t0veos", yname="tc_lo_b_t0veos_adj")

dequant2(tc_lo_b_eosvpod1, tc_lo_b_eosvpod1_adj)
de_euler(x=tc_lo_b_eosvpod1, y=tc_lo_b_eosvpod1_adj, xname="tc_lo_b_eosvpod1", yname="tc_lo_b_eosvpod1_adj")

dequant2(tc_lo_b_t0vpod1, tc_lo_b_t0vpod1_adj)
de_euler(x=tc_lo_b_t0vpod1, y=tc_lo_b_t0vpod1_adj, xname="tc_lo_b_t0vpod1", yname="tc_lo_b_t0vpod1_adj")

# Timecourse in high CRP group and treatment group A
dequant2(tc_hi_a_t0veos, tc_hi_a_t0veos_adj)
de_euler(x=tc_hi_a_t0veos, y=tc_hi_a_t0veos_adj, xname="tc_hi_a_t0veos", yname="tc_hi_a_t0veos_adj")

dequant2(tc_hi_a_eosvpod1, tc_hi_a_eosvpod1_adj)
de_euler(x=tc_hi_a_eosvpod1, y=tc_hi_a_eosvpod1_adj, xname="tc_hi_a_eosvpod1", yname="tc_hi_a_eosvpod1_adj")

dequant2(tc_hi_a_t0vpod1, tc_hi_a_t0vpod1_adj)
de_euler(x=tc_hi_a_t0vpod1, y=tc_hi_a_t0vpod1_adj, xname="tc_hi_a_t0vpod1", yname="tc_hi_a_t0vpod1_adj")

#### Timecourse in high CRP group and treatment group B
dequant2(tc_hi_b_t0veos, tc_hi_b_t0veos_adj)
de_euler(x=tc_hi_b_t0veos, y=tc_hi_b_t0veos_adj, xname="tc_hi_b_t0veos", yname="tc_hi_b_t0veos_adj")

dequant2(tc_hi_b_eosvpod1, tc_hi_b_eosvpod1_adj)
de_euler(x=tc_hi_b_eosvpod1, y=tc_hi_b_eosvpod1_adj, xname="tc_hi_b_eosvpod1", yname="tc_hi_b_eosvpod1_adj")

dequant2(tc_hi_b_t0vpod1, tc_hi_b_t0vpod1_adj)
de_euler(x=tc_hi_b_t0vpod1, y=tc_hi_b_t0vpod1_adj, xname="tc_hi_b_t0vpod1", yname="tc_hi_b_t0vpod1_adj")

```


### low CRP group

Few changes in T0 vs EOS and more in EOS vs POD1.

TODO:
* Euler diagram
* scatter plot of stat values
* pathway analysis
* bar charts and meaning of the DGE results

```{r,lowcrpquant}

dequant2(tc_lo_t0veos_adj, tc_lo_eosvpod1_adj)
de_euler(x=tc_lo_t0veos_adj, y=tc_lo_eosvpod1_adj, xname="tc_lo_t0veos_adj", yname="tc_lo_eosvpod1_adj")

```

### high CRP group

```{r,hicrpquant}

dequant2(tc_hi_t0veos_adj, tc_hi_eosvpod1_adj)
de_euler(x=tc_hi_t0veos_adj, y=tc_hi_eosvpod1_adj, xname="tc_hi_t0veos_adj", yname="tc_hi_eosvpod1_adj")

```

### Timecourse in low CRP group and treatment group A

```{r,lowcrp_a_quant}

dequant2(tc_lo_a_t0veos_adj, tc_lo_a_eosvpod1_adj)
de_euler(x=tc_lo_a_t0veos_adj, y=tc_lo_a_eosvpod1_adj, xname="tc_lo_a_t0veos_adj", yname="tc_lo_a_eosvpod1_adj")

```

### Timecourse in low CRP group and treatment group B

```{r,lowcrp_b_quant}

dequant2(tc_lo_b_t0veos_adj, tc_lo_b_eosvpod1_adj)
de_euler(x=tc_lo_b_t0veos_adj, y=tc_lo_b_eosvpod1_adj, xname="tc_lo_b_t0veos_adj", yname="tc_lo_b_eosvpod1_adj")

```

### Timecourse in high CRP group and treatment group A

```{r,hicrp_a_quant}

dequant2(tc_hi_a_t0veos_adj, tc_hi_a_eosvpod1_adj)
de_euler(x=tc_hi_a_t0veos_adj, y=tc_hi_a_eosvpod1_adj, xname="tc_hi_a_t0veos_adj", yname="tc_hi_a_eosvpod1_adj")

```

### Timecourse in high CRP group and treatment group B

```{r,hicrp_b_quant}

dequant2(tc_hi_b_t0veos_adj, tc_hi_b_eosvpod1_adj)
de_euler(x=tc_hi_b_t0veos_adj, y=tc_hi_b_eosvpod1_adj, xname="tc_hi_b_t0veos_adj", yname="tc_hi_b_eosvpod1_adj")

```

### Comparison of high and low CRP groups T0 vs EOS

```{r,t0veos_hilo}

dequant2(tc_lo_t0veos_adj, tc_hi_t0veos_adj)
de_euler(x=tc_lo_t0veos_adj, y=tc_hi_t0veos_adj, xname="tc_lo_t0veos_adj", yname="tc_hi_t0veos_adj")

```

### Comparison of high and low CRP groups EOS vs POD1

```{r,eosvpod1_hilo}

dequant2(tc_lo_eosvpod1_adj, tc_hi_eosvpod1_adj)
de_euler(x=tc_lo_eosvpod1_adj, y=tc_hi_eosvpod1_adj, xname="tc_lo_eosvpod1_adj", yname="tc_hi_eosvpod1_adj")

```

### Comparison of high and low CRP groups T0 vs POD1

```{r,tovpod1_hilo}

dequant2(tc_lo_t0vpod1_adj, tc_hi_t0vpod1_adj)
de_euler(x=tc_lo_t0vpod1_adj, y=tc_hi_t0vpod1_adj, xname="tc_lo_t0vpod1_adj", yname="tc_hi_t0vpod1_adj")

```

### Comparison of treatment groups A and B in low CRP T0 vs EOS 

```{r,lo_t0veos_avb}

dequant2(tc_lo_a_t0veos_adj, tc_lo_b_t0veos_adj)
de_euler(x=tc_lo_a_t0veos_adj, y=tc_lo_b_t0veos_adj, xname="tc_lo_a_t0veos_adj", yname="tc_lo_b_t0veos_adj")

```

### Comparison of treatment groups A and B in low CRP EOS vs POD1

```{r,lo_eosvpod1_avb}

dequant2(tc_lo_a_eosvpod1_adj, tc_lo_b_eosvpod1_adj)
de_euler(x=tc_lo_a_eosvpod1_adj, y=tc_lo_b_eosvpod1_adj, xname="tc_lo_a_eosvpod1_adj", yname="tc_lo_b_eosvpod1_adj")

```

### Comparison of treatment groups A and B in low CRP T0 vs POD1

```{r,lo_t0vpod1_avb}

dequant2(tc_lo_a_t0vpod1_adj, tc_lo_b_t0vpod1_adj)
de_euler(x=tc_lo_a_t0vpod1_adj, y=tc_lo_b_t0vpod1_adj, xname="tc_lo_a_t0vpod1_adj", yname="tc_lo_b_t0vpod1_adj")

```

### Comparison of treatment groups A and B in high CRP T0 vs EOS 

```{r,hi_t0veos_avb}

dequant2(tc_hi_a_t0veos_adj, tc_hi_b_t0veos_adj)
de_euler(x=tc_hi_a_t0veos_adj, y=tc_hi_b_t0veos_adj, xname="tc_hi_a_t0veos_adj", yname="tc_hi_b_t0veos_adj")

```

### Comparison of treatment groups A and B in high CRP EOS vs POD1

```{r,hi_eosvpod1_avb}

dequant2(tc_hi_a_eosvpod1_adj, tc_hi_b_eosvpod1_adj)
de_euler(x=tc_hi_a_eosvpod1_adj, y=tc_hi_b_eosvpod1_adj, xname="tc_hi_a_eosvpod1_adj", yname="tc_hi_b_eosvpod1_adj")

```

### Comparison of treatment groups A and B in high CRP T0 vs POD1

```{r,hi_t0vpod1_avb}

dequant2(tc_hi_a_t0vpod1_adj, tc_hi_b_t0vpod1_adj)
de_euler(x=tc_hi_a_t0vpod1_adj, y=tc_hi_b_t0vpod1_adj, xname="tc_hi_a_t0vpod1_adj", yname="tc_hi_b_t0vpod1_adj")

```

## STOP HERE

This one is getting long, so I will continue downstream analysis in a separate script.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
