---
title: "PADDI RNA expression analysis - Timecourse Analysis - Downstream"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Time course analysis of PADDI genomics data.

```{r,libraries}

suppressPackageStartupMessages({
  library("gplots")
  library("reshape2")
  library("WGCNA")
  library("dplyr")
  library("DESeq2")
  library("mitch")
  library("MASS")
  library("eulerr")
  library("kableExtra")
  library("gplots")
})

```

## Load the data

```{r,load}

load("tca_pairwise.Rdata")

go <- mitch::gmt_import("c5.go.v2024.1.Hs.symbols.gmt")
names(go) <- gsub("_"," ",names(go))

gt <- read.table("../ref/gencode.v38.genetable.tsv")

rownames(gt) <- paste(gt[,1],gt[,2])

gt[,1] <- rownames(gt)

```

## Overview

We will run timecourse analysis with a simple pairwise approach.
For each of the groups below, this will involve three comparisons:

* T0 vs EOS

* EOS vs POD1

* T0 vs POD1

The groups/subgroups we will look at are:

1. Timecourse in low CRP group

2. Timecourse in high CRP group

3. Timecourse in low CRP group and treatment group A

4. Timecourse in low CRP group and treatment group B

5. Timecourse in high CRP group and treatment group A

6. Timecourse in high CRP group and treatment group B

In this report I will be loading in the DESeq2 objects and conducting downstream analysis.

## Functions

```{r,mitchpwfuncs}

mitch1d <- function(de,dename,gs,gt=gt) {
  m <- mitch_import(x=de, DEtype="deseq2", geneTable=gt )
  mres <- mitch_calc(x=m,genesets=gs,minsetsize=5,cores=8,priority="effect")
  mtop_up <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05  & s.dist>0),15)
  mtop_up |> kbl(caption=paste(dename,"GO up")) |> kable_paper("hover", full_width = F)
  mtop_dn <- head(subset (mres$enrichment_result,p.adjustANOVA<0.05  & s.dist<0),15)
  mtop_dn |> kbl(caption=paste(dename,"GO dn")) |> kable_paper("hover", full_width = F)
  mtop <- rbind(mtop_up, mtop_dn[nrow(mtop_dn):1,])
  vec <- mtop$s.dist
  names(vec) <- mtop$set
  vec <- sort(vec)
  par( mar = c(5.1, 25.1, 4.1, 2.1) )
  barplot(vec,horiz=TRUE,las=1,cex.names=0.7,xlab="ES",main=dename)
  par( mar = c(5.1, 4.1, 4.1, 2.1) )
#  mitch_report(res=mres,outfile=paste(dename,"_mitchreport.html",sep=""),overwrite=TRUE)
}

#mitch1d(de=tc_lo_t0veos_adj, dename="tc_lo_t0veos_adj", gs=go, gt=gt)

mitch2d <- function(de1, de2, de1name, de2name, gs, gt=gt, ngenes=40) {
  l <-list(de1,de2)
  names(l) <- c(de1name, de2name)
  m <- mitch_import(x=l, DEtype="deseq2", geneTable=gt )
  mres <- mitch_calc(x=m,genesets=gs, minsetsize=5, cores=8, priority="effect")
  mtop <- head( subset(mres$enrichment_result, p.adjustMANOVA < 0.05), ngenes)
  mtop |> kbl(caption=paste("GO")) |> kable_paper("hover", full_width = F)
  mx <- as.matrix(mtop[,4:5])
  rownames(mx) <- mtop$set
  mx <- rbind(mx,t(as.matrix(c(-1,1))))
  rownames(mx)[nrow(mx)] <- "minmax"
  colfunc <- colorRampPalette(c("darkblue","blue","lightblue", "white", "pink","red","darkred"))
  heatmap.2(mx,col=colfunc(50),scale="none",trace="none",mar=c(8,20),cexRow=0.6, cexCol=0.8)
}

#mitch2d(de=tc_lo_t0veos_adj, de2=tc_lo_eosvpod1_adj,
#  de1name="tc_lo_t0veos_adj", de2name="tc_lo_eosvpod1_adj", gs=go, gt=gt)

```

### Object names

#### Timecourse in low CRP group

tc_lo_t0veos_adj
tc_lo_eosvpod1_adj
tc_lo_t0vpod1_adj

```{r,quant1}

mitch1d(de=tc_lo_t0veos_adj, dename="tc_lo_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_eosvpod1_adj, dename="tc_lo_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_t0vpod1_adj, dename="tc_lo_t0vpod1_adj", gs=go, gt=gt)

```

#### Timecourse in high CRP group

tc_hi_t0veos_adj
tc_hi_eosvpod1_adj
tc_hi_t0vpod1_adj

```{r,quant2}

mitch1d(de=tc_hi_t0veos_adj, dename="tc_hi_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_eosvpod1_adj, dename="tc_hi_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_t0vpod1_adj, dename="tc_hi_t0vpod1_adj", gs=go, gt=gt)

```

#### Timecourse in low CRP group and treatment group A

tc_lo_a_t0veos_adj
tc_lo_a_eosvpod1_adj
tc_lo_a_t0vpod1_adj

```{r,quant3}

mitch1d(de=tc_lo_a_t0veos_adj, dename="tc_lo_a_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_a_eosvpod1_adj, dename="tc_lo_a_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_a_t0vpod1_adj, dename="tc_lo_a_t0vpod1_adj", gs=go, gt=gt)

```

#### Timecourse in low CRP group and treatment group B

tc_lo_b_t0veos_adj
tc_lo_b_eosvpod1_adj
tc_lo_b_t0vpod1_adj

```{r,quant4}

mitch1d(de=tc_lo_b_t0veos_adj, dename="tc_lo_b_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_b_eosvpod1_adj, dename="tc_lo_b_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_lo_b_t0vpod1_adj, dename="tc_lo_b_t0vpod1_adj", gs=go, gt=gt)

```


#### Timecourse in high CRP group and treatment group A

tc_hi_a_t0veos_adj
tc_hi_a_eosvpod1_adj
tc_hi_a_t0vpod1_adj

```{r,quant5}

mitch1d(de=tc_hi_a_t0veos_adj, dename="tc_hi_a_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_a_eosvpod1_adj, dename="tc_hi_a_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_a_t0vpod1_adj, dename="tc_hi_a_t0vpod1_adj", gs=go, gt=gt)

```

#### Timecourse in high CRP group and treatment group B

tc_hi_b_t0veos_adj
tc_hi_b_eosvpod1_adj
tc_hi_b_t0vpod1_adj

```{r,quant6}

mitch1d(de=tc_hi_b_t0veos_adj, dename="tc_hi_b_t0veos_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_b_eosvpod1_adj, dename="tc_hi_b_eosvpod1_adj", gs=go, gt=gt)

mitch1d(de=tc_hi_b_t0vpod1_adj, dename="tc_hi_b_t0vpod1_adj", gs=go, gt=gt)

```

### low CRP group

Few changes in T0 vs EOS and more in EOS vs POD1.

TODO:
* Euler diagram
* scatter plot of stat values
* pathway analysis
* bar charts and meaning of the DGE results

```{r,lowcrpquant}

mitch2d(de1=tc_lo_t0veos_adj, de2=tc_lo_eosvpod1_adj,
  de1name="tc_lo_t0veos_adj", de2name="tc_lo_eosvpod1_adj", gs=go, gt=gt)

```

### high CRP group

```{r,hicrpquant}

mitch2d(de1=tc_hi_t0veos_adj, de2=tc_hi_eosvpod1_adj,
  de1name="tc_hi_t0veos_adj", de2name="tc_hi_eosvpod1_adj", gs=go, gt=gt)

```

### Timecourse in low CRP group and treatment group A

```{r,lowcrp_a_quant}

mitch2d(de1=tc_lo_a_t0veos_adj, de2=tc_lo_a_eosvpod1_adj,
  de1name="tc_lo_a_t0veos_adj", de2name="tc_lo_a_eosvpod1_adj", gs=go, gt=gt)

```

### Timecourse in low CRP group and treatment group B

```{r,lowcrp_b_quant}

mitch2d(de=tc_lo_b_t0veos_adj, de2=tc_lo_b_eosvpod1_adj,
  de1name="tc_lo_b_t0veos_adj", de2name="tc_lo_b_eosvpod1_adj", gs=go, gt=gt)

```

### Timecourse in high CRP group and treatment group A

```{r,hicrp_a_quant}

mitch2d(de1=tc_hi_a_t0veos_adj, de2=tc_hi_a_eosvpod1_adj,
  de1name="tc_hi_a_t0veos_adj", de2name="tc_hi_a_eosvpod1_adj", gs=go, gt=gt)

```

### Timecourse in high CRP group and treatment group B

```{r,hicrp_b_quant}

mitch2d(de1=tc_hi_b_t0veos_adj, de2=tc_hi_b_eosvpod1_adj,
  de1name="tc_hi_b_t0veos_adj", de2name="tc_hi_b_eosvpod1_adj", gs=go, gt=gt)

```

### Comparison of high and low CRP groups T0 vs EOS

```{r,t0veos_hilo}

mitch2d(de1=tc_lo_t0veos_adj, de2=tc_hi_t0veos_adj,
  de1name="tc_lo_t0veos_adj", de2name="tc_hi_t0veos_adj", gs=go, gt=gt)

```

### Comparison of high and low CRP groups EOS vs POD1

```{r,eosvpod1_hilo}

mitch2d(de1=tc_lo_eosvpod1_adj, de2=tc_hi_eosvpod1_adj ,
  de1name="tc_lo_eosvpod1_adj", de2name="tc_hi_eosvpod1_adj", gs=go, gt=gt)

```

### Comparison of high and low CRP groups T0 vs POD1

```{r,tovpod1_hilo}

mitch2d(de1=tc_lo_t0vpod1_adj, de2=tc_hi_t0vpod1_adj,
  de1name="tc_lo_t0vpod1_adj", de2name="tc_hi_t0vpod1_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in low CRP T0 vs EOS 

```{r,lo_t0veos_avb}

mitch2d(de1=tc_lo_a_t0veos_adj, de2=tc_lo_b_t0veos_adj,
  de1name="tc_lo_a_t0veos_adj", de2name="tc_lo_b_t0veos_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in low CRP EOS vs POD1

```{r,lo_eosvpod1_avb}

mitch2d(de1=tc_lo_a_eosvpod1_adj, de2=tc_lo_b_eosvpod1_adj,
  de1name="tc_lo_a_eosvpod1_adj", de2name="tc_lo_b_eosvpod1_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in low CRP T0 vs POD1

```{r,lo_t0vpod1_avb}

mitch2d(de1=tc_lo_a_t0vpod1_adj, de2=tc_lo_b_t0vpod1_adj,
  de1name="tc_lo_a_t0vpod1_adj", de2name="tc_lo_b_t0vpod1_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in high CRP T0 vs EOS 

```{r,hi_t0veos_avb}

mitch2d(de1=tc_hi_a_t0veos_adj, de2=tc_hi_b_t0veos_adj,
  de1name="tc_hi_a_t0veos_adj", de2name="tc_hi_b_t0veos_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in high CRP EOS vs POD1

```{r,hi_eosvpod1_avb}

mitch2d(de1=tc_hi_a_eosvpod1_adj, de2=tc_hi_b_eosvpod1_adj,
  de1name="tc_hi_a_eosvpod1_adj", de2name="tc_hi_b_eosvpod1_adj", gs=go, gt=gt)

```

### Comparison of treatment groups A and B in high CRP T0 vs POD1

```{r,hi_t0vpod1_avb}

mitch2d(de1=tc_hi_a_t0vpod1_adj, de2=tc_hi_b_t0vpod1_adj,
  de1name="tc_hi_a_t0vpod1_adj", de2name="tc_hi_b_t0vpod1_adj", gs=go, gt=gt)

```

## STOP HERE

This one is getting long, so I will continue downstream analysis in a separate script.

## Session information

For reproducibility

```{r,sessioninfo}

sessionInfo()

```
